---
title: "Lobster Exploration"
author: "Andre Dextre"
date: "2022-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(here)
```

Read in the data

```{r}
lobsters <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>%
  clean_names() %>%
  uncount(count)
```

- Create a summary table that finds the total counts (see: n()), and mean carapace lengths of lobsters observed in the dataset by site and year.

- Create a ggplot graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped (either aesthetically or by faceting) by site

```{r}
sum_table <- lobsters %>%
  group_by(site, year) %>%
  summarize(mean_length = mean(size_mm),
            count = n())

ggplot(sum_table, aes(x = year,
                      y = count)) +
  geom_line(aes(color = site)) 
```

The legal lobster size (carapace length) in California is 79.76 mm.

Create a subset that only contains lobster data from 2020 (note: this should be from the original data you read in, not the summary table you created above)

Write code (you can decide how to do this - there are a number of ways) to find the counts of lobsters observed at each site (only using site as the grouping factor) that are above and below the legal limit. Hint: You may want to add a new column legal that contains “yes” or “no” based on the size of the observed lobster (see dplyr::case_when() for a really nice way to do this), then use group_by() %>% summarize(n()) or dplyr::count() to get counts by group within variables

Create a stacked column graph that shows the proportion of legal and non-legal lobsters at each site. **Hint: create a stacked column graph with geom_col(), then add the argument position = "fill" to convert from a graph of absolute counts to proportions.

Which two sites had the largest proportion of legal lobsters in 2020? Explore the metadata to come up with a hypothesis about why that might be.

```{r}
lobsters_2020 <- lobsters %>%
  filter(year == 2020) %>%
  mutate(legal = case_when(size_mm  >= 79.76 ~ 'Yes',
            TRUE ~ 'No')) %>%
  group_by(site, legal) %>%
  summarize(counts = n()) %>%
  relocate(legal, .after = counts)

ggplot(lobsters_2020, aes(x = site,
                          y = counts,
                          group = legal)) +
  geom_col(position = "fill",
    aes(fill = legal)) +
  labs(title = "Proportion of Lobsters by CA Legality",
       x = "Site",
       y = "Count")
```


## Task 3: Random Lobster Wrangling

### Filter() practice

Create and store a subset that only contains lobsters from sites “IVEE”, “CARP” and “NAPL”. Check your output data frame to ensure that only those three sites exist.

```{r}
ex_a <- lobsters %>%
  filter(site %in% c("IVEE", "CARP", "NAPL"))
```

Create a subset that only contains lobsters observed in August.

```{r}
ex_b <- lobsters %>%
  filter(month == 8)
```

Create a subset with lobsters at Arroyo Quemado (AQUE) OR with a carapace length greater than 70 mm.

```{r}
ex_c <- lobsters %>%
  filter(site == "AQUE" | size_mm > 70)
```

Create a subset that does NOT include observations from Naples Reef (NAPL)

```{r}
ex_d <- lobsters %>%
  filter(site != "NAPL")
```

### group_by() %>% summarize() practice

Find the mean and standard deviation of lobster carapace length, grouped by site.

```{r}
ex_e <- lobsters %>%
  group_by(site) %>%
  summarize(mean_mm = mean(size_mm, na.rm = TRUE),
            sd_mm = sd(size_mm, na.rm = TRUE))
```

Find the maximum carapace length by site and month.

```{r}
ex_f <- lobsters %>%
  group_by(site, month) %>%
  summarize(max = max(size_mm, na.rm = TRUE))
```

### `mutate()` practice

Add a new column that contains lobster carapace length converted to centimeters. Check output. 

```{r}
ex_g <- lobsters %>%
  mutate(length_cm = size_mm / 10)
```

Update the site column to all lowercase. Check output.

```{r}
ex_h <- lobsters %>%
  mutate(site = tolower(site))
```

Convert the area column to a character (not sure why you’d want to do this, but try it anyway). Check output

```{r}
ex_i <- lobsters %>%
  mutate(area = as.character(area))
```


